R code for 'Apparent physiological costs of blood parasite infection in the early-life of a vertebrate host - only in highest infections

### 2022-02-23 MS_physiology_infection.Rmd
Add files via upload

```{r packages, message=F, warning=F}
library(chron)
library(cowplot)
library(EnvStats)
library(lme4)
library(lsmeans)
library(ggalt)
library(ggdist)
library(ggeffects)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(ggridges)
library(ggsci)
library(ggsignif)
library(grid)
library(gridExtra)
library(magrittr)
library(patchwork)
library(plyr)
library(rptR)
library(RVAideMemoire)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(tidybayes)
library(tidyverse)
library(wesanderson)
theme_set(theme_tidybayes() + panel_border())
```

@@

```{r data}
## Load data from csv file and code variable classes
## -----------------------------------------------------------------------------
df <- 
  read.csv('./physiology_infection_stage_full_wideformat.csv', h = T, sep = ';', dec = ".") %>% 
  mutate(Nest = as.factor(Nest)) %>% 
  mutate(Date = chron(as.character(Date), format = "d/m/y")) %>% 
  mutate(Infection_intensity.1 = as.factor(Infection_intensity.1)) %>% 
  mutate(Infection_intensity.2 = as.factor(Infection_intensity.2)) %>%
  mutate(Prevalence2 = as.factor(ifelse(Infection_intensity.2 == 0, 0, 1))) %>%
  mutate(Prevalence1 = as.factor(ifelse(Infection_intensity.1 == 0, 0, 1))) %>% 
  mutate(Breathing_rate = as.numeric(Breathing_rate)) %>% 
  mutate(Sex = as.factor(Sex)) %>% 
  mutate(Sampling_rank = as.factor(Sampling_rank)) %>% 
  mutate(Weight_g = scale(as.numeric(Weight.2))) %>% 
  mutate(Wing_cm = as.numeric(Wing.2)) %>% 
  mutate(body_temperature = as.numeric(body_temperature)) %>% 
  mutate(delta_body_condition = as.numeric(delta_body_condition)) %>% 
  mutate(Ambient_Temperature_AvgC = scale(as.numeric(TemperatureAvgC))) %>% 
  mutate(Treatment = as.factor(Treatment)) %>% 
  mutate(Year = as.factor(Year)) %>% 
  mutate(Growth_rate = as.numeric(Growth_rate)) %>% 
  mutate(days_resample = scale(days_resample)) %>% 
  mutate(Age_in_days = scale(as.numeric(Age_in_days))) %>% 
  mutate(Diff_Inf_Class = factor(Diff_Inf_Class, levels = c("Uninfected","Increase","Peak","Decrease")))

@@

```{r}
## Build contingency table: Infection prevalence and treatment: First sampling
## -----------------------------------------------------------------------------
pre_treatment <- with(df, table(Prevalence1, Treatment))
prop.test(pre_treatment)
pairwise.prop.test(pre_treatment)

@@

## Plotting count of infected / uninfected nestlings across treatments
## -----------------------------------------------------------------------------
tab_cont = data.frame(
  Count = c(infected_treatment[1,], uninfected_treatment[1,], infected_treatment[2,], uninfected_treatment[2,]),
  Timepoint = rep(c("Pre-treatment","Post-treatment"), 4),
  Treatment = rep(c("Control", "Treatment"), each = 4),
  Prevalence = rep(rep(c("Infected", "Uninfected"), each = 2),2)) %>% 
  mutate(Timepoint = factor(Timepoint, levels = c("Pre-treatment","Post-treatment"))) %>% 
  mutate(Prevalence = factor(Prevalence, levels = c("Uninfected","Infected")))

@@

## Plotting mean infection intensity values of nestlings across treatments
## -----------------------------------------------------------------------------
mean_infint = data.frame(
  Sampling_rank = factor(c(rep("1",length(df$Sampling_rank)),as.character(df$Sampling_rank)),levels = c("1","2"),
                         labels  =  c("Pre-treatment","Post-treatment")),
  Infection_intensity = as.numeric(c(as.character(df$Infection_intensity.1),as.character(df$Infection_intensity.2))),
  Treatment = factor(c(as.character(rep(df$Treatment,2))),levels  =  c("Control","Malarone")))

## Figure 2b
## -----------------------------------------------------------------------------
treatment_meaninf_changes = ggplot(mean_infint) +
  aes(Sampling_rank, Infection_intensity, group = Treatment) +
  geom_vline(xintercept  =  "Pre-treatment", linetype = "dashed", color = 'darkgrey') +
  geom_vline(xintercept = "Post-treatment", linetype = "dashed", color = 'darkgrey') +
  labs(y = "Mean infection intensity") +
  geom_point(aes(colour = Treatment),position = position_jitterdodge(jitter.height  =  0.1))+
  stat_summary(fun = mean, mapping = aes(group=Treatment), geom = "line",size = 2,color = c("lightgrey","black","black","lightgrey"))+ 
  stat_summary(fun = mean, geom = "point",size = 5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.04, size = 2) + 
  scale_color_manual(values = c("lightgrey","black")) + 
  theme_classic() + 
  theme(axis.title.x = element_blank(),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16),
        axis.title.y = element_text(size = 18, margin =  unit(c(0, 3, 0, 0), "mm")),
        axis.text = element_text(size = 16),
        plot.margin = unit(c(0,0,0,30), "pt"))

@@

## Testing for difference in mean infection intensity between Pre and Post treatment
## -----------------------------------------------------------------------------
mean_before_after <- data.frame(
  Ring_no = 
    c(df$Ring_no[df$Sampling_rank %in% 2], df$Ring_no[df$Sampling_rank %in% 2]),
  Infection_intensity = c(as.character(df$Infection_intensity.1), as.character(df$Infection_intensity.2)),
  Treatment = 
    c(as.character(df$Treatment[df$Sampling_rank %in% 2]), as.character(df$Treatment[df$Sampling_rank %in% 2])),
  Timepoint = 
    c(rep("Pre-treatment", times = length(df$Treatment)), rep("Post-treatment", times = length(df$Infection_intensity.2)))) %>% 
  mutate(Treatment = as.factor(Treatment)) %>% 
  mutate(Timepoint = as.factor(Timepoint))

mal = mean_before_after[mean_before_after$Treatment %in% "Malarone",]
ctrl = mean_before_after[mean_before_after$Treatment %in% "Control",]
first = mean_before_after[mean_before_after$Timepoint %in% "Pre-treatment",]

## Pre-treatment difference in parasitic load between Treatments
## -----------------------------------------------------------------------------
perm.t.test(as.numeric(as.character(first$Infection_intensity)) ~ first$Treatment, progress = F)
sd(first$Infection_intensity[first$Treatment %in% "Control"])

## In malarone treated nestlings
## -----------------------------------------------------------------------------
perm.t.test(as.numeric(as.character(mal$Infection_intensity)) ~ mal$Timepoint, progress = F)

# In control nestlings
## -----------------------------------------------------------------------------
perm.t.test(as.numeric(as.character(ctrl$Infection_intensity))~ctrl$Timepoint, progress = F)
```

@@

## Modeling host health to Infection intensity
## -----------------------------------------------------------------------------
tolres_mod=lm(delta_body_condition~Infection_intensity.2, data=tol_resdata)

>>>>

```{r packages, message=F, warning=F}
library(chron)
library(cowplot)
library(EnvStats)
library(lme4)
library(lsmeans)
library(ggalt)
library(ggdist)
library(ggeffects)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(ggridges)
library(ggsci)
library(ggsignif)
library(grid)
library(gridExtra)
library(magrittr)
library(patchwork)
library(plyr)
library(rptR)
library(RVAideMemoire)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(tidybayes)
library(tidyverse)
library(wesanderson)
library(car)
theme_set(theme_tidybayes() + panel_border())
```

@@

```{r data}
## Load data from csv file and code variable classes
## -----------------------------------------------------------------------------
df <- 
  read.csv('./physiology_infection_stage_full_wideformat.csv', h = T, sep = ';', dec = ".") %>% 
  mutate(Nest = as.factor(Nest)) %>% 
  mutate(Date = chron(as.character(Date), format = "d/m/y")) %>% 
  mutate(Infection_intensity.1 = as.factor(Infection_intensity.1)) %>% 
  mutate(Infection_intensity.2 = as.factor(Infection_intensity.2)) %>%
  mutate(Prevalence2 = as.factor(ifelse(Infection_intensity.2 == 0, 0, 1))) %>%
  mutate(Prevalence1 = as.factor(ifelse(Infection_intensity.1 == 0, 0, 1))) %>% 
  mutate(Breathing_rate = as.numeric(Breathing_rate)) %>% 
  mutate(Sex = as.factor(Sex)) %>% 
  mutate(Sampling_rank = as.factor(Sampling_rank)) %>% 
  mutate(Weight_g = scale(as.numeric(Weight.2))) %>% 
  mutate(Wing_cm = as.numeric(Wing.2)) %>% 
  mutate(body_temperature = as.numeric(body_temperature)) %>% 
  mutate(delta_body_condition = as.numeric(delta_body_condition)) %>% 
  mutate(Ambient_Temperature_AvgC = scale(as.numeric(TemperatureAvgC))) %>% 
  mutate(Treatment = as.factor(Treatment)) %>% 
  mutate(Year = as.factor(Year)) %>% 
  mutate(Growth_rate = as.numeric(Growth_rate)) %>% 
  mutate(days_resample = scale(days_resample)) %>% 
  mutate(Age_in_days = scale(as.numeric(Age_in_days))) %>% 
  mutate(intensity_change = as.numeric(intensity_change)) %>% 
  mutate(Diff_Inf_Class = factor(Diff_Inf_Class, levels = c("Uninfected","Increase","Peak","Decrease")))

@@

```{r, results='hide', message=F, warning=F,prompt=F}

## Build contingency table: Infection prevalence and treatment: First sampling
## -----------------------------------------------------------------------------
pre_treatment <- with(df, table(Prevalence1, Treatment))
prop.test(pre_treatment)
pairwise.prop.test(pre_treatment)

@@

## Plotting count of infected / uninfected nestlings across treatments
## -----------------------------------------------------------------------------
tab_cont = data.frame(
  Count = c(infected_treatment[1,], uninfected_treatment[1,], infected_treatment[2,], uninfected_treatment[2,]),
  Timepoint = rep(c("First sampling","Second sampling"), 4),
  Treatment = rep(c("Control", "Treatment"), each = 4),
  Prevalence = rep(rep(c("Infected", "Uninfected"), each = 2),2)) %>% 
  mutate(Timepoint = factor(Timepoint, levels = c("First sampling","Second sampling"))) %>% 
  mutate(Prevalence = factor(Prevalence, levels = c("Uninfected","Infected")))

@@

## Plotting mean infection intensity values of nestlings across treatments
## -----------------------------------------------------------------------------
mean_infint = data.frame(
  Sampling_rank = factor(c(rep("1",length(df$Sampling_rank)),as.character(df$Sampling_rank)),levels = c("1","2"),
                         labels  =  c("First sampling","Second sampling")),
  Infection_intensity = as.numeric(c(as.character(df$Infection_intensity.1),as.character(df$Infection_intensity.2))),
  Treatment = factor(c(as.character(rep(df$Treatment,2))),levels  =  c("Control","Malarone")))

## Figure 2b
## -----------------------------------------------------------------------------
treatment_meaninf_changes = ggplot(mean_infint) +
  aes(Sampling_rank, Infection_intensity, group = Treatment) +
  geom_vline(xintercept  =  "First sampling", linetype = "dashed", color = 'darkgrey') +
  geom_vline(xintercept = "Second sampling", linetype = "dashed", color = 'darkgrey') +
  labs(y = "Mean infection intensity") +
  geom_point(aes(colour = Treatment),position = position_jitterdodge(jitter.height  =  0.1))+
  stat_summary(fun = mean, geom = "line",size = 2,color = c("lightgrey","black","black","lightgrey"))+ 
  stat_summary(fun = mean, geom = "point",size = 5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.04, size = 2) + 
  scale_color_manual(values = c("lightgrey","black")) + 
  theme_classic() + 
  theme(axis.title.x = element_blank(),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16),
        axis.title.y = element_text(size = 18, margin =  unit(c(0, 3, 0, 0), "mm")),
        axis.text = element_text(size = 16),
        plot.margin = unit(c(0,0,0,30), "pt"))

@@

## Testing for difference in parasitic load between Treatment groups at first sampling
## -----------------------------------------------------------------------------

perm.t.test(as.numeric(as.character(df$Infection_intensity.1))~df$Treatment)

## Testing for Treatment effect on parasitemia change between first and second sampling
## -----------------------------------------------------------------------------

lmerTest::lmer(intensity_change~Treatment+Year+days_resample+(1|Nest), data=df[!df$intensity_change %in% -100,]) %>%
  summary()
```

@@

## Modeling host health to Infection intensity
## -----------------------------------------------------------------------------

tolres_mod=lm(delta_body_condition~Infection_intensity.2, data=tol_resdata)
